`localhierarchy`: an R package to facilitate fitting of global and local Bayesian hierarchical models

### 2025-10-08 paper/paper.qmd
paper updates

---
title: ""
author: ""
format:
  pdf:
    latex-engine: pdflatex    # lualatex or xelatex or pdflatex
    include-before-body: title.tex
    include-in-header: 
      - file: preamble.tex 
editor: visual
execute:
  cache: true
abstract: | 
  Bayesian hierarchical models are widely used in global health estimation, where data availability may vary across national or subnational populations. Such models are typically fitted to a global database, e.g., to produce national-level estimates for all countries in the world or in some region. To facilitate analysis at a local level, models are often desirable that are informed by global models but can be fitted to just a subset of the data, such as data for one country. We refer to such models as local models: models that are derived from global hierarchical models but adapted such that they can be fitted to the data from one population alone.

@@

![Data, estimates, and forecasts from 1970 until 2030 for FP indicators for married women in Benin (BEN), Gambia (GMB), and Ghana (GHA). Graphs show mCPR, unmet need for modern methods, demand, and demand satisfied with modern methods. Survey data are shown in colored dots with 95% confidence intervals that reflect total error variance. FPET estimates are shown in black, with grey shaded areas representing 90% credible intervals.](fig/fp_local.jpg){#fig-localfits}

<!-- #| label: fig-localfits -->

<!-- #| fig-cap: "Data, estimates, and forecasts from 1970 until 2030 for FP indicators for married women in Benin (BEN), Gambia (GMB), and Ghana (GHA). Graphs show mCPR, unmet need for modern methods, demand, and demand satisfied with modern methods. Survey data are shown in colored dots with 95% confidence intervals that reflect total error variance. FPET estimates are shown in black, with grey shaded areas representing 90% credible intervals." -->

<!-- plots <- readRDS("outputfpet_localfits.rds") -->

<!-- plots[[1]] /  plots[[2]]/ plots[[3]]  -->

@@

Given limited country-specific information on some parameters, Bayesian hierarchical models are used to share information between populations, using hierarchical structures such as the one illustrated in Figure \ref{fig-hierstruc}. Specifically for family planning estimation and forecasting, a hierarchical model is used to estimate the level of demand in a reference year and the asymptote of the transition of demand from low to high values. @fig-level shows estimates of the level in the reference year at the country level for countries in Western Africa, and for the groupings used in the hierarchical model. @fig-asympt shows estimates for the asymptote. These figures show variation in levels among countries in Western Africa, informed by country-specific data. Estimates of asymptotes are similar across countries in Western Africa because countries have not yet reached the end phase of their transition; the asymptotes in Western African countries are informed by the estimate of the asymptote in sub-Saharan Africa through the hierarchical model.

@@

```{=latex}
\begin{figure}[h]
\begin{adjustbox}{max size={\textwidth}{\textheight},center}
%\resizebox{\textwidth}{\textwidth}{
\begin{tikzpicture}[%scale=0.8,
  box/.style={rectangle, draw, rounded corners, minimum width=2.6cm, minimum height=1cm, align=center},
  level0/.style={box, fill=world!40},
  level1/.style={box, fill=blue!20},
  level2/.style={box, fill=purple!20},
  level3/.style={box, fill=country!40, minimum width=2.2cm},
    level4/.style={box, fill=data!20, minimum width=2.2cm},
  arrow/.style={-{Latex[length=3mm]}, thick},
  faintarrow/.style={-{Latex[length=2mm]}, thick, opacity=0.4},
  background/.style 2 args={rectangle, rounded corners, draw=none, fill=#1, fill opacity=0.15, inner sep=0.5cm, fit=#2}
]

% --- Level 0 ---
\node[level0] (mu0) {World};

% --- Level 1 ---
\node[level1, below left=1cm and 3cm of mu0] (mu11) {Sub-Saharan Africa};
\node[level1, right=of mu11] (mu12) {South-East Asia};
\node[level1, right=of mu12] (dots1) {$\dots$};
% --- Level 2 ---
\node[level2, below left=1cm and 0cm of mu11] (mu21) {Western Africa};
\node[level2, right=of mu21] (mu22) {Eastern Africa};
\node[level2, right=of mu22] (dots2) {$\dots$};
% % --- Level 3 ---
\node[level3, below left=1cm and 0cm of mu21] (mu31) {Senegal};
\node[level3, right=of mu31] (mu32) {Burkina Faso};
\node[level3, right=of mu32] (dots3) {$\dots$};
% % % --- Level 4 ---
% \node[level4, below left=1cm and 0cm of mu31] (mu41) {Data point 1};
% \node[level4, right=of mu41] (mu42) {Data point 2};
% \node[level4, right=of mu42] (dots4) {$\dots$};
% --- Background blocks ---
\begin{pgfonlayer}{background}
  \node[background={world!40}{(mu0)}] {};
  \node[background={blue!20}{(mu11)(mu12)(dots1)}] (bg1) {};
  \node[background={pink!60}{(mu21)(mu22)(dots2) }] (bg2) {};
 \node[background={country!40}{(mu31)(mu32)(dots3)}] (bg3) {};
 % \node[background={data!40}{(mu41)(mu42)(dots4)}] (bg4) {};
\end{pgfonlayer}

% Sigma labels inside background blocks
\node[anchor=north west] at (bg1.east) {Major world regions};
\node[anchor=north west] at (bg2.east) {Smaller world regions};
\node[anchor=north west] at (bg3.east) {Countries};
%\node[anchor=north west] at (bg4.east) %{Country data};
% --- Arrows ---
\draw[arrow] (mu0.south) -- (mu11.north);
\draw[arrow] (mu0.south) -- (mu12.north);
\draw[arrow] (mu0.south) -- (dots1.north);

% \draw[arrow] (mu31.south) -- (mu41.north);
% \draw[arrow] (mu31.south) -- (mu42.north);
% \draw[arrow] (mu31.south) -- (dots4.north);
\end{tikzpicture}%}
\end{adjustbox}
\caption{Illustration of hierarchical grouping of countries in world regions.}\label{fig-hierstruc}
\end{figure}
```

<!-- ```{r, echo = FALSE} -->

<!-- plots <- readRDS("outputfpet_params.rds") -->

@@

<!-- #| label: fig-level -->

<!-- #| fig-cap: "Estimates of the level of demand for countries in Western Africa and for higher levels used in the hierarchical model (smaller world regions and major world regions). Point estimates and 95% credible intervals are shown. " -->

<!-- plots$plot_reg_omega[[2]]/ plots$plot_reg_omega[[3]]/  plots$plot_iso_omega -->

<!-- #| label: fig-asympt -->

<!-- #| fig-cap: "Estimates of the asymptote for demand for countries in Western Africa and for higher levels used in the hierarchical model (major world regions). Point estimates and 95% credible intervals are shown. " -->

<!-- plots$plot_reg_ptilde[[2]]  / plots$plot_iso_ptilde -->

@@

Suppose we want to estimate a population parameter $\mu_r^{(l)}$ of interest, referring to a population health indicator for region $r$ at level $l$ of the hierarchy. For example, the indicator could refer to some population health indicator such as service coverage or morbidity or mortality-related outcomes. For a given level $l$, index $r$ refers to the population in a particular region indexed by $r$, for example, a subnational region, a country, or a group of countries. We assume that $\mu$ is unconstrained (possibly after a transformation has been applied) and consider the following general hierarchical model setup, see Figure \ref{fig-hiermodel}:

@@

@@

We use the term local hierarchical models to refer to a hierarchical model in which a subset of parameters have been fixed at posterior mean estimates obtained from a fit of the model to a larger data set. A typical workflow for local estimation of national level outcomes is illustrated in Figure \ref{fig-workflow}. For national estimation using the model given by Eq.\ref{eq-countryhier}, the user starts by fitting a global model, using a global data base and estimating all parameters in a hierarchical model. In a local model, data from one country are used and parameters that are not country-specific are fixed using parameter estimates from the global model fit.

```{=latex}
\begin{figure}[h]
\begin{adjustbox}{max size={\textwidth}{\textheight},center}
\begin{tikzpicture}[
  box/.style={rectangle, draw, thick, rounded corners, minimum width=6cm, align=left, inner sep=6pt},
  databox/.style={box, fill=data!20},
  modelbox/.style={box, fill=param!20},
  arrowdata/.style={-{Latex[length=3mm]}, thick, data},
  arrowparam/.style={-{Latex[length=3mm]}, thick, param!80},
  font=\sffamily
]
% --- Nodes ---
\node[databox] (globaldata) {Global data base of country data};
\node[modelbox, below=1cm of globaldata] (globalmodel) {%
  Global national modeling:\\
  Estimate all parameters in hierarchical model, e.g., \\
\(
  \begin{aligned}
\mu_w^{(region)} &\sim N(\mu^{(global)}, \sigma_{ region}^2),\\
\mu_s^{(subregion)} &\sim N(\mu_{w[s]}^{(region)}, \sigma_{ subregion}^2), \\
\mu_c^{(country)} &\sim N(\mu_{s[c]}^{(subregion)}, \sigma_{ country}^2),  \end{aligned}
\)
};
\node[modelbox, below=1cm of globalmodel] (localmodel) {%
  Local national modeling:\\
  Estimate national parameters only, \\
  fix parameter estimates at higher levels of the hierarchy \\i.e., $\mu_c^{(country)}$ with $\mu_c^{(country)} \sim N(\hat{\mu}_{s[c]}^{(subregion)}, \hat{\sigma}_{ country}^2).$
};

% --- Data arrow (down then right) ---
\draw[arrowdata] (globaldata.south) -- (globalmodel.north);
\draw[arrowdata] (globaldata.east) ++(0,0) -- ++(4.5,0) |- (localmodel.east)
node[midway,above] {Data for one country};

% --- Parameter arrow ---
\draw[arrowparam] (globalmodel.south) -- (localmodel.north);

% --- Legend ---
% \node[arrowparam, below=1.2cm of localmodel.west, anchor=west] (leg1) {}; 
% \node[right=0.3cm of leg1] {\textcolor{param}{$\rightarrow$ Parameter estimates}};
\node[below=1cm of localmodel](parameter) {\textcolor{param}{$\rightarrow$ Parameter estimates}};
\node[right=1cm of parameter] {\textcolor{data}{$\rightarrow$ Data}};
\end{tikzpicture}
\end{adjustbox}
\caption{Workflow for global and local national modeling. }\label{fig-workflow}
\end{figure}
```

@@

The simulation setup is summarized in Figure \ref{fig-simusetup}, top and center panels. Simulation of subnational data is discussed in a later section.

```{=latex}
\begin{figure}[h]
\fbox{\begin{adjustbox}{max size={\textwidth}{\textheight},center}
\begin{tikzpicture}[
  box/.style={rectangle, draw, rounded corners, minimum width=2cm, minimum height=1cm, align=center},
  level0/.style={box, fill=yellow!40},
  level1/.style={box, fill=orange!40},
  level2/.style={box, fill=green!20},
  arrow/.style={-{Latex[length=3mm]}, thick},
  faintarrow/.style={-{Latex[length=2mm]}, thick, opacity=0.4},
  background/.style 2 args={rectangle, rounded corners, draw=none, fill=#1, fill opacity=0.15, inner sep=0.5cm, fit=#2}
]

% --- Level 0 ---
\node[level0] (mu0) {$\mu^{(\glob)} \sim \mathcal{N}(0,s_\glob^2)$};

% --- Level 1 ---
\node[level1, below left=1.5cm and 3cm of mu0] (mu11) {$\mu^{(\country)}_1 \sim \mathcal{N}(\mu^{(\glob)},\sigma_{\country}^2)$};
\node[level1, right=of mu11] (mu12) {$\mu^{(\country)}_2$};
\node[level1, right=of mu12] (dots1) {$\dots$};
% --- Level 2 ---
\node[level2, below left=1.5cm and 0cm of mu11] (mu21) {$\mu^{(\subnat)}_1 \sim \mathcal{N}(\mu^{(\country)}_1,\sigma_{\subnat}^2)$};
\node[level2, right=of mu21] (mu22) {$\mu^{(2)}_2$};
\node[level2, right=of mu22] (dots2) {$\dots$};
% --- Background blocks ---
\begin{pgfonlayer}{background}
  \node[background={yellow!40}{(mu0)}] {};
  \node[background={orange!40}{(mu11)(mu12)(dots1)}] (bg1) {};
  \node[background={green!40}{(mu21)(mu22)(dots2) }] (bg2) {};
\end{pgfonlayer}

% Sigma labels inside background blocks
\node[anchor=north west] at (bg1.east) {$\sigma_{\country}$, country};
\node[anchor=north west] at (bg2.east) {$\sigma_{\subnat}$, subnational region};
% --- Arrows ---
\draw[arrow] (mu0.south) -- (mu11.north);
\draw[arrow] (mu0.south) -- (mu12.north);
\draw[arrow] (mu0.south) -- (dots1.north);

\draw[arrow] (mu11.south) -- (mu21.north);
\draw[arrow] (mu11.south) -- (mu22.north);
\draw[arrow] (mu11.south) -- (dots2.north);
\end{tikzpicture}
\end{adjustbox}}

% local model nat
\fbox{\begin{adjustbox}{max size={\textwidth}{\textheight},center}
\begin{tikzpicture}[
  box/.style={rectangle, draw, rounded corners, minimum width=2cm, minimum height=1cm, align=center},
  level0/.style={box, fill=yellow!40},
  level1/.style={box, fill=orange!40},
  level2/.style={box, fill=green!20},
    data/.style={box, fill=gray!20},
  arrow/.style={-{Latex[length=3mm]}, thick},
  faintarrow/.style={-{Latex[length=2mm]}, thick, opacity=0.4},
  background/.style 2 args={rectangle, rounded corners, draw=none, fill=#1, fill opacity=0.15, inner sep=0.5cm, fit=#2}
]

\node[level1] (mu0) {$\mu^{(\country)}_1$};

% --- Level 1 ---
\node[data, below left=1.5cm and 3cm of mu0] (mu11) {$y^{(country)}_1 \sim \mathcal{N}(\mu^{(\country)}_1,\sigma_{y}^2)$};
\node[data, right=of mu11] (mu12) {$y^{(country)}_2$};
\node[data, right=of mu12] (dots1) {$\dots$};
% --- Background blocks ---
\begin{pgfonlayer}{background}
  \node[background={gray!40}{(mu11)(mu12)(dots1)}] (bg1) {};
\end{pgfonlayer}

% Sigma labels inside background blocks
\node[anchor=north west] at (bg1.east) {$\sigma_{y}$, country data for first country};

% --- Arrows ---
\draw[arrow] (mu0.south) -- (mu11.north);
\draw[arrow] (mu0.south) -- (mu12.north);
\draw[arrow] (mu0.south) -- (dots1.north);
\end{tikzpicture}
\end{adjustbox}}

% local model subnat
\fbox{
\begin{adjustbox}{max size={\textwidth}{\textheight},center}
\begin{tikzpicture}[
  box/.style={rectangle, draw, rounded corners, minimum width=2cm, minimum height=1cm, align=center},
  level0/.style={box, fill=yellow!40},
  level1/.style={box, fill=orange!40},
  level2/.style={box, fill=green!20},
    data/.style={box, fill=gray!20},
  arrow/.style={-{Latex[length=3mm]}, thick},
  faintarrow/.style={-{Latex[length=2mm]}, thick, opacity=0.4},
  background/.style 2 args={rectangle, rounded corners, draw=none, fill=#1, fill opacity=0.15, inner sep=0.5cm, fit=#2}
]

% keep same colors so now color of level 1 at top 
\node[level2] (mu0) {$\mu^{(\subnat)}_1$};

% --- Level 1 ---
\node[data, below left=1.5cm and 3cm of mu0] (mu11) {$y^{(subnat)}_1 \sim \mathcal{N}(\mu^{(\subnat)}_1,\sigma_{y}^2)$};
\node[data, right=of mu11] (mu12) {$y^{(subnat)}_2$};
\node[data, right=of mu12] (dots1) {$\dots$};
% --- Background blocks ---
\begin{pgfonlayer}{background}
  \node[background={gray!40}{(mu11)(mu12)(dots1)}] (bg1) {};
\end{pgfonlayer}

% Sigma labels inside background blocks
\node[anchor=north west] at (bg1.east) {$\sigma_{y}$, subnational data for first region};

% --- Arrows ---
\draw[arrow] (mu0.south) -- (mu11.north);
\draw[arrow] (mu0.south) -- (mu12.north);
\draw[arrow] (mu0.south) -- (dots1.north);
\end{tikzpicture}
\end{adjustbox}}
\caption{Diagram of simulation settings. Top: Hierarchical model for generating country and subnational parameters. Center: Data generating mechanism for national data, illustrated for the first country. Bottom: Data generating mechanism for subnational data, illustrated for the first subnational region (used only in Section 5).}\label{fig-simusetup}
\end{figure}
\clearpage
```

>>>>

---
title: ""
author: ""
format:
  pdf:
    latex-engine: pdflatex    # lualatex or xelatex or pdflatex
    include-before-body: title.tex
    include-in-header: 
      - file: preamble.tex 
editor: visual
abstract: | 
  Bayesian hierarchical models are widely used in global health estimation, where data availability may vary across national or subnational populations. Such models are typically fitted to a global database, e.g., to produce national-level estimates for all countries in the world or in some region. To facilitate analysis at a local level, models are often desirable that are informed by global models but can be fitted to just a subset of the data, such as data for one country. We refer to such models as local models: models that are derived from global hierarchical models but adapted such that they can be fitted to the data from one population alone.

@@

![Data, estimates, and forecasts from 1970 until 2030 for FP indicators for married women in Benin (BEN), Gambia (GMB), and Ghana (GHA). Graphs show modern contraceptive use, unmet need for modern methods, demand, and demand satisfied with modern methods. Survey data are shown in colored dots with 95% confidence intervals that reflect total error variance. FPET estimates are shown in black, with grey shaded areas representing 90% credible intervals.](fig/fp_local.jpg){#fig-localfits}

@@

Given limited country-specific information on some parameters, Bayesian hierarchical models are used to share information between populations, using hierarchical structures such as the one illustrated in \@fig-hierstruc. Specifically for family planning estimation and forecasting, a hierarchical model is used to estimate the level of demand in a reference year and the asymptote of the transition of demand from low to high values. @fig-level shows estimates of the level in the reference year at the country level for countries in Western Africa, and for the groupings used in the hierarchical model. @fig-asympt shows estimates for the asymptote. These figures show variation in levels among countries in Western Africa, informed by country-specific data. Estimates of asymptotes are similar across countries in Western Africa because countries have not yet reached the end phase of their transition; the asymptotes in Western African countries are informed by the estimate of the asymptote in sub-Saharan Africa through the hierarchical model.

@@

<!-- ```{=latex} -->

<!-- \begin{figure}[h] -->

<!-- \begin{adjustbox}{max size={\textwidth}{\textheight},center} -->

<!-- %\resizebox{\textwidth}{\textwidth}{ -->

<!-- \begin{tikzpicture}[%scale=0.8, -->

<!--   box/.style={rectangle, draw, rounded corners, minimum width=2.6cm, minimum height=1cm, align=center}, -->

<!--   level0/.style={box, fill=world!40}, -->

<!--   level1/.style={box, fill=blue!20}, -->

<!--   level2/.style={box, fill=purple!20}, -->

<!--   level3/.style={box, fill=country!40, minimum width=2.2cm}, -->

<!--     level4/.style={box, fill=data!20, minimum width=2.2cm}, -->

<!--   arrow/.style={-{Latex[length=3mm]}, thick}, -->

<!--   faintarrow/.style={-{Latex[length=2mm]}, thick, opacity=0.4}, -->

<!--   background/.style 2 args={rectangle, rounded corners, draw=none, fill=#1, fill opacity=0.15, inner sep=0.5cm, fit=#2} -->

<!-- ] -->

<!-- % --- Level 0 --- -->

<!-- \node[level0] (mu0) {World}; -->

<!-- % --- Level 1 --- -->

<!-- \node[level1, below left=1cm and 3cm of mu0] (mu11) {Sub-Saharan Africa}; -->

<!-- \node[level1, right=of mu11] (mu12) {South-East Asia}; -->

<!-- \node[level1, right=of mu12] (dots1) {$\dots$}; -->

<!-- % --- Level 2 --- -->

<!-- \node[level2, below left=1cm and 0cm of mu11] (mu21) {Western Africa}; -->

<!-- \node[level2, right=of mu21] (mu22) {Eastern Africa}; -->

<!-- \node[level2, right=of mu22] (dots2) {$\dots$}; -->

<!-- % % --- Level 3 --- -->

<!-- \node[level3, below left=1cm and 0cm of mu21] (mu31) {Senegal}; -->

<!-- \node[level3, right=of mu31] (mu32) {Burkina Faso}; -->

<!-- \node[level3, right=of mu32] (dots3) {$\dots$}; -->

<!-- % % % --- Level 4 --- -->

<!-- % \node[level4, below left=1cm and 0cm of mu31] (mu41) {Data point 1}; -->

<!-- % \node[level4, right=of mu41] (mu42) {Data point 2}; -->

<!-- % \node[level4, right=of mu42] (dots4) {$\dots$}; -->

<!-- % --- Background blocks --- -->

<!-- \begin{pgfonlayer}{background} -->

<!--   \node[background={world!40}{(mu0)}] {}; -->

<!--   \node[background={blue!20}{(mu11)(mu12)(dots1)}] (bg1) {}; -->

<!--   \node[background={pink!60}{(mu21)(mu22)(dots2) }] (bg2) {}; -->

<!--  \node[background={country!40}{(mu31)(mu32)(dots3)}] (bg3) {}; -->

<!--  % \node[background={data!40}{(mu41)(mu42)(dots4)}] (bg4) {}; -->

<!-- \end{pgfonlayer} -->

<!-- % Sigma labels inside background blocks -->

<!-- \node[anchor=north west] at (bg1.east) {Major world regions}; -->

<!-- \node[anchor=north west] at (bg2.east) {Smaller world regions}; -->

<!-- \node[anchor=north west] at (bg3.east) {Countries}; -->

<!-- %\node[anchor=north west] at (bg4.east) %{Country data}; -->

<!-- % --- Arrows --- -->

<!-- \draw[arrow] (mu0.south) -- (mu11.north); -->

<!-- \draw[arrow] (mu0.south) -- (mu12.north); -->

<!-- \draw[arrow] (mu0.south) -- (dots1.north); -->

<!-- \draw[arrow] (mu11.south) -- (mu21.north); -->

<!-- \draw[arrow] (mu11.south) -- (mu22.north); -->

<!-- \draw[arrow] (mu11.south) -- (dots2.north); -->

<!-- \draw[arrow] (mu21.south) -- (mu31.north); -->

<!-- \draw[arrow] (mu21.south) -- (mu32.north); -->

<!-- \draw[arrow] (mu21.south) -- (dots3.north); -->

<!-- % \draw[arrow] (mu31.south) -- (mu41.north); -->

<!-- % \draw[arrow] (mu31.south) -- (mu42.north); -->

<!-- % \draw[arrow] (mu31.south) -- (dots4.north); -->

<!-- \end{tikzpicture}%} -->

<!-- \end{adjustbox} -->

<!-- \caption{Illustration of hierarchical grouping of countries in world regions.}\label{fig-hierstruc} -->

<!-- \end{figure} -->

@@

![Illustration of hierarchical grouping of countries in world regions.](fig/worldregions.jpg){#fig-hierstruc width="643"}

@@

Suppose we want to estimate a population parameter $\mu_r^{(l)}$ of interest, referring to a population health indicator for region $r$ at level $l$ of the hierarchy. For example, the indicator could refer to some population health indicator such as service coverage or morbidity or mortality-related outcomes. For a given level $l$, index $r$ refers to the population in a particular region indexed by $r$, for example, a subnational region, a country, or a group of countries. We assume that $\mu$ is unconstrained (possibly after a transformation has been applied) and consider the following general hierarchical model setup, see @fig-hiermodel:

@@

![General specification of a hierarchical model.](fig/hiermodel.jpg){#fig-hiermodel width="664"}

@@

We use the term local hierarchical models to refer to a hierarchical model in which a subset of parameters have been fixed at posterior mean estimates obtained from a fit of the model to a larger data set. A typical workflow for local estimation of national level outcomes is illustrated in @fig-workflow. For national estimation using the model given by Eq.\ref{eq-countryhier}, the user starts by fitting a global model, using a global data base and estimating all parameters in a hierarchical model. In a local model, data from one country are used and parameters that are not country-specific are fixed using parameter estimates from the global model fit.

<!-- ```{=latex} -->

<!-- \begin{figure}[h] -->

<!-- \begin{adjustbox}{max size={\textwidth}{\textheight},center} -->

<!-- \begin{tikzpicture}[ -->

<!--   box/.style={rectangle, draw, thick, rounded corners, minimum width=6cm, align=left, inner sep=6pt}, -->

<!--   databox/.style={box, fill=data!20}, -->

<!--   modelbox/.style={box, fill=param!20}, -->

<!--   arrowdata/.style={-{Latex[length=3mm]}, thick, data}, -->

<!--   arrowparam/.style={-{Latex[length=3mm]}, thick, param!80}, -->

<!--   font=\sffamily -->

<!-- ] -->

<!-- % --- Nodes --- -->

<!-- \node[databox] (globaldata) {Global data base of country data}; -->

<!-- \node[modelbox, below=1cm of globaldata] (globalmodel) {% -->

<!--   Global national modeling:\\ -->

<!--   Estimate all parameters in hierarchical model, e.g., \\ -->

<!-- \( -->

<!--   \begin{aligned} -->

<!-- \mu_w^{(region)} &\sim N(\mu^{(global)}, \sigma_{ region}^2),\\ -->

<!-- \mu_s^{(subregion)} &\sim N(\mu_{w[s]}^{(region)}, \sigma_{ subregion}^2), \\ -->

<!-- \mu_c^{(country)} &\sim N(\mu_{s[c]}^{(subregion)}, \sigma_{ country}^2),  \end{aligned} -->

<!-- \) -->

<!-- }; -->

<!-- \node[modelbox, below=1cm of globalmodel] (localmodel) {% -->

<!--   Local national modeling:\\ -->

<!--   Estimate national parameters only, \\ -->

<!--   fix parameter estimates at higher levels of the hierarchy \\i.e., $\mu_c^{(country)}$ with $\mu_c^{(country)} \sim N(\hat{\mu}_{s[c]}^{(subregion)}, \hat{\sigma}_{ country}^2).$ -->

<!-- }; -->

<!-- % --- Data arrow (down then right) --- -->

<!-- \draw[arrowdata] (globaldata.south) -- (globalmodel.north); -->

<!-- \draw[arrowdata] (globaldata.east) ++(0,0) -- ++(4.5,0) |- (localmodel.east) -->

<!-- node[midway,above] {Data for one country}; -->

<!-- % --- Parameter arrow --- -->

<!-- \draw[arrowparam] (globalmodel.south) -- (localmodel.north); -->

<!-- % --- Legend --- -->

<!-- % \node[arrowparam, below=1.2cm of localmodel.west, anchor=west] (leg1) {};  -->

<!-- % \node[right=0.3cm of leg1] {\textcolor{param}{$\rightarrow$ Parameter estimates}}; -->

<!-- \node[below=1cm of localmodel](parameter) {\textcolor{param}{$\rightarrow$ Parameter estimates}}; -->

<!-- \node[right=1cm of parameter] {\textcolor{data}{$\rightarrow$ Data}}; -->

<!-- \end{tikzpicture} -->

<!-- \end{adjustbox} -->

<!-- \caption{Workflow for global and local national modeling. }\label{fig-workflow} -->

<!-- \end{figure} -->

![Workflow for global and local national modeling.](fig/workflow.jpg){#fig-workflow width="677"}

@@

The simulation setup is summarized in @fig-simusetup, top and center panels. Simulation of subnational data is discussed in a later section.

<!-- ```{=latex} -->

<!-- \begin{figure}[h] -->

<!-- \fbox{\begin{adjustbox}{max size={\textwidth}{\textheight},center} -->

<!-- \begin{tikzpicture}[ -->

<!--   box/.style={rectangle, draw, rounded corners, minimum width=2cm, minimum height=1cm, align=center}, -->

<!--   level0/.style={box, fill=yellow!40}, -->

<!--   level1/.style={box, fill=orange!40}, -->

<!--   level2/.style={box, fill=green!20}, -->

<!--   arrow/.style={-{Latex[length=3mm]}, thick}, -->

<!--   faintarrow/.style={-{Latex[length=2mm]}, thick, opacity=0.4}, -->

<!--   background/.style 2 args={rectangle, rounded corners, draw=none, fill=#1, fill opacity=0.15, inner sep=0.5cm, fit=#2} -->

<!-- ] -->

<!-- % --- Level 0 --- -->

<!-- \node[level0] (mu0) {$\mu^{(\glob)} \sim \mathcal{N}(0,s_\glob^2)$}; -->

<!-- % --- Level 1 --- -->

<!-- \node[level1, below left=1.5cm and 3cm of mu0] (mu11) {$\mu^{(\country)}_1 \sim \mathcal{N}(\mu^{(\glob)},\sigma_{\country}^2)$}; -->

<!-- \node[level1, right=of mu11] (mu12) {$\mu^{(\country)}_2$}; -->

<!-- \node[level1, right=of mu12] (dots1) {$\dots$}; -->

<!-- % --- Level 2 --- -->

<!-- \node[level2, below left=1.5cm and 0cm of mu11] (mu21) {$\mu^{(\subnat)}_1 \sim \mathcal{N}(\mu^{(\country)}_1,\sigma_{\subnat}^2)$}; -->

<!-- \node[level2, right=of mu21] (mu22) {$\mu^{(2)}_2$}; -->

<!-- \node[level2, right=of mu22] (dots2) {$\dots$}; -->

<!-- % --- Background blocks --- -->

<!-- \begin{pgfonlayer}{background} -->

<!--   \node[background={yellow!40}{(mu0)}] {}; -->

<!--   \node[background={orange!40}{(mu11)(mu12)(dots1)}] (bg1) {}; -->

<!--   \node[background={green!40}{(mu21)(mu22)(dots2) }] (bg2) {}; -->

<!-- \end{pgfonlayer} -->

<!-- % Sigma labels inside background blocks -->

<!-- \node[anchor=north west] at (bg1.east) {$\sigma_{\country}$, country}; -->

<!-- \node[anchor=north west] at (bg2.east) {$\sigma_{\subnat}$, subnational region}; -->

<!-- % --- Arrows --- -->

<!-- \draw[arrow] (mu0.south) -- (mu11.north); -->

<!-- \draw[arrow] (mu0.south) -- (mu12.north); -->

<!-- \draw[arrow] (mu0.south) -- (dots1.north); -->

<!-- \draw[arrow] (mu11.south) -- (mu21.north); -->

<!-- \draw[arrow] (mu11.south) -- (mu22.north); -->

<!-- \draw[arrow] (mu11.south) -- (dots2.north); -->

<!-- \end{tikzpicture} -->

<!-- \end{adjustbox}} -->

<!-- % local model nat -->

<!-- \fbox{\begin{adjustbox}{max size={\textwidth}{\textheight},center} -->

<!-- \begin{tikzpicture}[ -->

<!--   box/.style={rectangle, draw, rounded corners, minimum width=2cm, minimum height=1cm, align=center}, -->

<!--   level0/.style={box, fill=yellow!40}, -->

<!--   level1/.style={box, fill=orange!40}, -->

<!--   level2/.style={box, fill=green!20}, -->

<!--     data/.style={box, fill=gray!20}, -->

<!--   arrow/.style={-{Latex[length=3mm]}, thick}, -->

<!--   faintarrow/.style={-{Latex[length=2mm]}, thick, opacity=0.4}, -->

<!--   background/.style 2 args={rectangle, rounded corners, draw=none, fill=#1, fill opacity=0.15, inner sep=0.5cm, fit=#2} -->

<!-- ] -->

<!-- \node[level1] (mu0) {$\mu^{(\country)}_1$}; -->

<!-- % --- Level 1 --- -->

<!-- \node[data, below left=1.5cm and 3cm of mu0] (mu11) {$y^{(country)}_1 \sim \mathcal{N}(\mu^{(\country)}_1,\sigma_{y}^2)$}; -->

<!-- \node[data, right=of mu11] (mu12) {$y^{(country)}_2$}; -->

<!-- \node[data, right=of mu12] (dots1) {$\dots$}; -->

<!-- % --- Background blocks --- -->

<!-- \begin{pgfonlayer}{background} -->

<!--   \node[background={gray!40}{(mu11)(mu12)(dots1)}] (bg1) {}; -->

<!-- \end{pgfonlayer} -->

<!-- % Sigma labels inside background blocks -->

<!-- \node[anchor=north west] at (bg1.east) {$\sigma_{y}$, country data for first country}; -->

<!-- % --- Arrows --- -->

<!-- \draw[arrow] (mu0.south) -- (mu11.north); -->

<!-- \draw[arrow] (mu0.south) -- (mu12.north); -->

<!-- \draw[arrow] (mu0.south) -- (dots1.north); -->

<!-- \end{tikzpicture} -->

<!-- \end{adjustbox}} -->

<!-- % local model subnat -->

<!-- \fbox{ -->

<!-- \begin{adjustbox}{max size={\textwidth}{\textheight},center} -->

<!-- \begin{tikzpicture}[ -->

<!--   box/.style={rectangle, draw, rounded corners, minimum width=2cm, minimum height=1cm, align=center}, -->

<!--   level0/.style={box, fill=yellow!40}, -->

<!--   level1/.style={box, fill=orange!40}, -->

<!--   level2/.style={box, fill=green!20}, -->

<!--     data/.style={box, fill=gray!20}, -->

<!--   arrow/.style={-{Latex[length=3mm]}, thick}, -->

<!--   faintarrow/.style={-{Latex[length=2mm]}, thick, opacity=0.4}, -->

<!--   background/.style 2 args={rectangle, rounded corners, draw=none, fill=#1, fill opacity=0.15, inner sep=0.5cm, fit=#2} -->

<!-- ] -->

<!-- % keep same colors so now color of level 1 at top  -->

<!-- \node[level2] (mu0) {$\mu^{(\subnat)}_1$}; -->

<!-- % --- Level 1 --- -->

<!-- \node[data, below left=1.5cm and 3cm of mu0] (mu11) {$y^{(subnat)}_1 \sim \mathcal{N}(\mu^{(\subnat)}_1,\sigma_{y}^2)$}; -->

<!-- \node[data, right=of mu11] (mu12) {$y^{(subnat)}_2$}; -->

<!-- \node[data, right=of mu12] (dots1) {$\dots$}; -->

<!-- % --- Background blocks --- -->

<!-- \begin{pgfonlayer}{background} -->

<!--   \node[background={gray!40}{(mu11)(mu12)(dots1)}] (bg1) {}; -->

<!-- \end{pgfonlayer} -->

<!-- % Sigma labels inside background blocks -->

<!-- \node[anchor=north west] at (bg1.east) {$\sigma_{y}$, subnational data for first region}; -->

<!-- % --- Arrows --- -->

<!-- \draw[arrow] (mu0.south) -- (mu11.north); -->

<!-- \draw[arrow] (mu0.south) -- (mu12.north); -->

<!-- \draw[arrow] (mu0.south) -- (dots1.north); -->

<!-- \end{tikzpicture} -->

<!-- \end{adjustbox}} -->

<!-- \caption{Diagram of simulation settings. Top: Hierarchical model for generating country and subnational parameters. Center: Data generating mechanism for national data, illustrated for the first country. Bottom: Data generating mechanism for subnational data, illustrated for the first subnational region (used only in Section 5).}\label{fig-simusetup} -->

<!-- \end{figure} -->

<!-- \clearpage -->

![Diagram of simulation settings. Top: Hierarchical model for generating country and subnational parameters. Center: Data generating mechanism for national data, illustrated for the first country. Bottom: Data generating mechanism for subnational data, illustrated for the first subnational region (used only in Section 5).](fig/simusetup.jpg){#fig-simusetup width="682"}
